\documentclass[11pt, a4paper, twoside]{report}

\usepackage[british]{babel}
\usepackage[useregional]{datetime2}
\DTMlangsetup[en-GB]{showdayofmonth=false} %to remove day from date
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{listings}
\usepackage{hyperref}
\hypersetup{colorlinks=false}
\usepackage{lscape} %to put things landscape
\usepackage{rotating} %to put things landscape, again
\usepackage{pdflscape} %trying landscape again
%\usepackage{subfigure} %removed as it was interfering with subfig package
\usepackage{amsmath} %for some maths stuff! 
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{csvsimple}
\usepackage{lscape}%for landscape table in sup mat
\usepackage{apalike} %for round brackets in references
\usepackage{titlesec} %for hiding chapter title
\usepackage{subfig} % for getting 4 figures looking nice
\usepackage{setspace} %for spacing
\usepackage{listings} %for my R code
\usepackage{tikz} %for doing little circles in line with text
%\onehalfspacing %one and a half line spacing
\setstretch{1.5}
\titleformat{\chapter}[display]
  {\normalfont\bfseries}{}{0pt}{\Huge}

%%%%%This stuff is for generating tables%%%%
\usepackage{booktabs} % For \toprule, \midrule and \bottomrule
\usepackage{siunitx} % Formats the units and values
\usepackage{pgfplotstable} % Generates table from .csv

% Setup siunitx:
\sisetup{
  round-mode          = places, % Rounds numbers
  round-precision     = 3, % to 2 places
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

\title{Is everything everywhere? Deterministic and stochastic processes in microbial biogeography}
\author{Amy Solman}
% Update supervisor and other title stuff in title/title.tex

\begin{document}
\input{title.tex}

\renewcommand{\abstractname}{Declaration}
\begin{abstract}
\textbf{Concept:} The concept for this work came from Prof. Thomas Bell and Prof. Ryan Chisholm.\\
\noindent \textbf{Data:} All data for this analysis was collected by myself from the literature as listed in the Supplementary Materials section.\\
\noindent \textbf{Simulation:} All simulation code was written by myself, excluding function 2: coalescence\_test (simulation scripts) provided by Dr. James Rosindell.\\
\noindent \textbf{Model:} The Classic, Depth and Perimeter models, as well as critical area formulas, were supplied by Prof. Ryan Chisholm.\\
\noindent \textbf{Analysis:} I declare that all analysis was carried out by myself.\\
\noindent \textbf{Report:} I declare that the report was written by myself.\\
\\ \\ \\

\noindent \textbf{COVID-19} \\
\noindent The hypotheses presented in this report were originally investigated using a laboratory-based experiment in Professor Thomas Bell's Microbial Ecology Laboratory, Imperial College London, Silwood Park. Two months into the investigation, laboratory work was ceased due to the COVID-19 pandemic. The following report seeks to test the same hypotheses, using data from the literature.
\end{abstract}

\renewcommand{\abstractname}{Abstract}
\begin{abstract}
1. Microbial communities play an essential role in biogeochemical processes and ecosystem functioning. Despite their importance, relatively little is known about the mechanisms driving spatial scaling within these communities. \\

\noindent 2. The Theory of Island Biogeography predicts that species richness increases with island area through stochastic colonisation-extinction processes. It has been widely use to describe macroorganism spatial patterns, with little application to microbial communities.\\

\noindent 3. Despite the popularity of this theory, small islands often show no clear relationship between species richness and area. Chisholm \textit{et al.,} have addressed this by developing a unified theory of species-area relationships that transitions from a niche-structured regime at smaller spatial scales to stochastic mechanisms at larger spatial scales.\\

\noindent 4. This work modifies the Chisholm model to address microbial communities in a range of habitats. Through model fitting I assess to what extent microorganisms are subject to deterministic or stochastic processes. I also explore how the transition between regimes differs among habitat types and taxonomic groups. \\

\noindent 5. The models gave good fits to the data. Multiple regression analysis showed significant evidence that taxonomic group explained variance in the critical area of transition between regimes. Less motile groups exhibited higher critical areas supporting my hypothesis. Habitat type was non-significant in predicting critical area.\\

\noindent 6. A proportion of the datasets exhibited a biphasic species-area relationship. Broadly the critical area hypotheses of lower transition areas for more motile taxa was supported by the data. These results can assist in predicting the spatial scaling of microbial diversity, with application to climate change modelling. \\

\noindent \textbf{Keywords:} species-area relationships, taxa-area relationships, microbial biogeography, small-island effect, niche structure, colonisation-extinction balance, island biogeography
\end{abstract}

\tableofcontents
\listoffigures
\listoftables

\input{introduction.tex}
\input{methods.tex}
\input{results.tex}
\input{discussion.tex}
\input{datacodeavailability.tex}
\input{acknowledgements.tex}

\bibliographystyle{apalike}
\bibliography{bibliography}
\addcontentsline{toc}{chapter}{Bibliography}

\input{supplementary.tex}
\end{document}